### 2014-02-11 msc-thesis/chapters/1/chapter-intro.tex
MSC danas edits.

% What is being addressed in this thesis
Clearly, indirect conflicts and its subsequent research areas have a large breadth of limitations, some of which
will be explored in this dissertation. The research goal of this dissertation is to explore the limitations which exist
with supporting indirect conflicts through awareness techniques as well as to explore possible solution for industry practice 
in the are of indirect conflicts. To accomplish this goal, I have researcher the following sub topics of indirect
conflicts: technical dependencies, how developers are involved in said dependencies,
socio-technical congruence as a mitigation strategy to indirect conflicts, what the root causes of indirect conflicts are,
what compounding factors exist for indirect conflicts, what current industry mitigation strategies of indirect conflicts
are being used, what future steps should be taken in academia to better industry regarding indirect conflicts, and finally,
how software evolution analysis can be used to better tools for indirect conflicts. I have addressed these issues by
conducting the following four studies.

@@

% Study 1 introduction
Study 1 focuses on the power of technical dependencies in software projects. The question I
investigated was : ``\textit{Is it possible to identify pairs of developers whose technical dependencies in 
code changes statistically relate to bugs?}''. This study explains the approach used to locate these pairs of developers 
in developer networks.The process utilizes code changes and the call hierarchies effected to find patterns of developer 
relationships in successful and failed code changes. These developers pairs can be seen as indirect conflicts occurring
as one developer's code change has negatively affects another developer's work. As it will be seen, I found 27 statistically significant failure 
inducing developer pairs. These developer relationships can be used to promote the idea of leveraging socio-technical 
congruence, a measure of coordination compared to technical dependencies amongst stakeholders, to provide coordination recommendations.
This notion of socio-technical congruence is my initial proposed solution to indirect conflicts. By identifying these failure
inducing pair of developers over indirect conflicts, I hoped that recommended communication would be the correct fix. The 
results of Study 1 directly influence Study 2.

@@

In order to fully understand the root causes of information overload, false positives, and
scalability issues in regards to indirect conflicts, Study 3 was an empirical study to determine what events occur to
cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. 
I also set out to understand what mitigation strategies developers currently use as opposed to those created
by academia. Through this exploration, I examined what can be accomplished moving forward with indirect conflicts
in both research and industry. This study asked the following 3 research questions: 
\textit{What events or conditions lead to indirect conflicts?}
\textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
\textit{What do developers want from future indirect conflict tools?}

>>>>

% What is being addressed in this thesis
Clearly, indirect conflicts and its subsequent research areas have a large breadth of limitations, some of which
will be explored in this dissertation. The research goal of this dissertation is to explore the limitations which exist
with supporting indirect conflicts through awareness techniques as well as to explore possible solution for industry practice 
in the are of indirect conflicts. To accomplish this goal, I have researched the following sub topics of indirect
conflicts: technical dependencies, how developers are involved in said dependencies,
socio-technical congruence as a mitigation strategy to indirect conflicts, what the root causes of indirect conflicts are,
what compounding factors exist for indirect conflicts, what current industry mitigation strategies of indirect conflicts
are being used, what future steps should be taken by researchers to better industry regarding indirect conflicts, and finally,
how software evolution analysis can be used to better tools for indirect conflicts. I have addressed these issues by
conducting the following four studies.

@@

% Study 1 introduction
Study 1 focuses on the power of technical dependencies in software projects. The question I
investigated were : ``\textit{Is it possible to identify pairs of developers whose technical dependencies in 
code changes statistically relate to bugs?}''. This study explains the approach used to locate these pairs of developers 
in developer networks.The process utilizes code changes and the call hierarchies effected to find patterns of developer 
relationships in successful and failed code changes. These developers pairs can be seen as indirect conflicts occurring
as one developer's code change has negatively affects another developer's work. As it will be seen, I found 27 statistically significant failure 
inducing developer pairs. These developer relationships can be used to promote the idea of leveraging socio-technical 
congruence, a measure of coordination compared to technical dependencies amongst stakeholders, to provide coordination recommendations.
This notion of socio-technical congruence is my initial proposed solution to indirect conflicts. By identifying these failure
inducing pair of developers over indirect conflicts, I hoped that recommended communication would be the correct fix. The 
results of Study 1 directly influence Study 2.

@@

In order to fully understand the root causes of information overload, false positives, and
scalability issues in regards to indirect conflicts, Study 3 was an empirical study to determine what events occur to
cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. 
I also set out to understand what mitigation strategies developers currently use as opposed to those created
by researchers. Through this exploration, I examined what can be accomplished moving forward with indirect conflicts
in both research and industry. This study asked the following 3 research questions: 
\textit{What are the types, factors, and frequencies of indirect conflicts?}
\textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
\textit{What do developers want from future indirect conflict tools?}

### 2014-02-11 msc-thesis/chapters/2/sec-developer-pairs.tex
MSC danas edits.

>>>>

\subsection{Related Work}

Research has shown multiple reasons for software failures in both technical dependencies as well
as human or social dependencies in software development. On the technical side, studies have shown that
technical dependencies in software are often powerful predictors or errors in software as well as
in builds~\cite{Pinzger:2008:DNP, Zimmermann:2008:PDU, Hassan:4019574}. These technical dependencies
are often accompanied by data mining algorithms in order to set apart failure inducing dependencies
from non failing dependencies.

On the other side with the human factor, researchers have examined predicting build outcomes of software
using communication patterns from developers. Wolf et al.~\cite{5070503} used patterns of communication from between
pre-existing build in order to predict later build outcomes. Naggappan et al.~\cite{nagappan08} showed that having a large
communication organization difference between developers who worked on the same software module had
a negative influence on the quality of the software.

Studies have also combined both technical and social dependencies into a notion of socio-technical congruence.
Cataldo et al.~\cite{Cataldo:2006:ICR} have shown that this notion of socio-technical congruence can be leveraged to predict
and improve task completion times in a software projects.

These studies have mostly focused at a very high and abstract level of software development (builds and task completion).
Where they have fallen short is in fine grained analysis of software changes. This study is used to take the
pre existing knowledge of both technical and social dependencies and apply it to a source code change level of
granularity. Instead of large scale failures like long completion times or build failures, this study examines failures
at the bug level induced by each code change.

### 2014-02-11 msc-thesis/chapters/2/sec-impact.tex
MSC danas edits.

I first start with both Codebook~\cite{Begel:2010:CDE} and 
Ariadne~\cite{Trainer:2005:BGT}. These projects produce an exploratory
environment for developers to handle indirect conflicts. Exploratory
pertains to the ability to solve self determined conflicts, meaning that
once a developer discovers a conflict, they can use the tool as a type of
lookup table to solve their issue. Codebook is a type of social developer
network that relates developers to source code, issue repositories and
other social media while Ariadne only looks at source code for developer
to source code association. Through Codebook, developers become
owners of source code artifacts. Both projects also use program 
dependency graphs~\cite{Horwitz:1992:UPD}
in order to relate technical artifacts to each other. These projects make 
use of method call graphs in order to 
determine which methods invoke others which forms the basis for 
linking source code artifacts creating a directed graph. While these 
projects can be great tools 
for solving indirect conflicts which may arise, by querying such directed
graphs to view impacts of conflict creating code, they lack the ability to
detect potential conflicts on their own.

@@

While technical object relationships form the basis of indirect conflicts,
communication between developers is my ultimate goal of resolving such conflicts
(as was seen in Study 1).
This being the case, developer ownership must be placed on the 
identified technical objects. With this ownership, we now infer
relationships among developers based on their technical objects
``uses'' relationship. Developer A, who owns object 1, which uses 
object 2 owned by developer B, may be notified by a change to
object 2's internal workings. Most, if not all, ownership information
of technical objects can be extracted from a project's source code
repository (CVS, Git, SVN, etc.).

@@

These two studies have shown that my design and approach to
detecting and alerting developers to indirect conflicts appear
to be on the correct path. However, \textit{Impact} has clearly shown
the achilles heel of indirect conflict tools, which is information
overload because of an inability to detect ``notification worthy''
changes.

\subsection{Threats to Validity}
% TODO

>>>>

I first start with both Codebook~\cite{Begel:2010:CDE} and 
Ariadne~\cite{Trainer:2005:BGT}. These projects produce an exploratory
environment for developers to handle indirect conflicts. Exploratory
pertains to the ability to solve self determined conflicts, meaning that
once a developer discovers a conflict, they can use the tool as a type of
lookup table to solve their issue. Codebook is a type of social developer
network that relates developers to source code, issue repositories and
other social media while Ariadne only examines there source code for developer
to source code association. Through Codebook, developers become
owners of source code artifacts. Both projects also use program 
dependency graphs~\cite{Horwitz:1992:UPD}
in order to relate technical artifacts to each other. These projects make 
use of method call graphs in order to 
determine which methods invoke others which forms the basis for 
linking source code artifacts creating a directed graph. While these 
projects can be great tools 
for solving indirect conflicts which may arise, by querying such directed
graphs to view impacts of conflict creating code, they lack the ability to
detect potential conflicts on their own.

@@

While technical object relationships form the basis of indirect conflicts,
communication between developers is my ultimate goal of resolving such conflicts
(as was seen in Study 1).
This being the case, developer ownership must be placed on the 
identified technical objects. With this ownership, I now infer
relationships among developers based on their technical objects
``uses'' relationship. Developer A, who owns object 1, which uses 
object 2 owned by developer B, may be notified by a change to
object 2's internal workings. Most, if not all, ownership information
of technical objects can be extracted from a project's source code
repository (CVS, Git, SVN, etc.).

@@

These two studies have shown that my design and approach to
detecting and alerting developers to indirect conflicts appear
to be on the correct path. However, \textit{Impact} has clearly shown
the Achilles heel of indirect conflict tools, which is information
overload because of an inability to detect ``notification worthy''
changes.

\subsection{Threats to Validity}
Because of the tool validation nature of this work, I chose participant interviews as my research validation method, which has some implications regarding the
limitations and threats to validity of this study. While I did some positive results regarding the potential of Impact in this study
populations studied from outside of this studies participants may add new insights into the pool of findings. As a result of this, findings
from this study may not relate to everyone or generalize to outside of the group of participants involved in this study.

As per my interview participants, I conducted this case study with undergraduate students at the University of Victoria. This being said, participants
may not have had enough real world experience to validate this study at an industry level. The students were also consumed with their regular course
work which could have limited the time spend using Impact and the enthusiasm put forward while conducting this study.

To counter act this, my study was conducted purely on a volunteer basis to eliminate those participants which may have been to busy to focus
on the study or provide appropriate feedback where needed.

### 2014-02-11 msc-thesis/chapters/3/chapter-exploration.tex
MSC danas edits.

While these issues of information overload, false positives, dependencies, and scalability continue to come up
in most indirect conflict tools, only a handful of attempts have been made at rectifying these issues or
finding the root causes~\cite{Holmes:2010:CAR,Kim:2011:ESA}. 
In order to fully understand the root causes of information overload, false positives, and
scalability issues in regards to indirect conflicts in this study, I examine and determine what events occur to
cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. By determining the root
causes of source code changes in indirect conflicts, we may be able to create indirect conflict tools which have filtered
monitoring in order to only detect those changes with a high likelihood of causing indirect conflicts.
I then determine what mitigation strategies developers currently use as opposed to those created
by academia. Since developers have identified indirect conflicts as a major concern for themselves, but at
the same time are not using the tools put forth by academia, I wish to find what they use to mitigate indirect
conflicts. Through these findings, we can create tools which are similar to those already in use by developers in
the hopes of a higher adoption rate of tools produced by academia.
Finally, I examine what can be accomplished moving forward with indirect conflicts
in both research and industry.

>>>>

While these issues of information overload, false positives, dependencies, and scalability continue to come up
in most indirect conflict tools, only a handful of attempts have been made at rectifying these issues or
finding the root causes~\cite{Holmes:2010:CAR,Kim:2011:ESA}. 
In order to fully understand the root causes of information overload, false positives, and
scalability issues in regards to indirect conflicts in this study, I examine and determine what events occur to
cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. By determining the root
causes of source code changes in indirect conflicts, we may be able to create indirect conflict tools which have filtered
monitoring in order to only detect those changes with a high likelihood of causing indirect conflicts.
I then determine what mitigation strategies developers currently use as opposed to those created
by researchers. Since developers have identified indirect conflicts as a major concern for themselves, but at
the same time are not using the tools put forth by researchers, I wish to find what they use to mitigate indirect
conflicts. Through these findings, we can create tools which are similar to those already in use by developers in
the hopes of a higher adoption rate of tools produced by researchers.
Finally, I examine what can be accomplished moving forward with indirect conflicts
in both research and industry.

### 2014-02-11 msc-thesis/chapters/3/sec-apie.tex
MSC danas edits.

Once the ChangeDistiller information was collected, I decided to examine software change trends surrounding releases of the project's I had selected. Since releases
have preconceived notions of software stability and a lack of proneness to indirect conflicts, 
I decided that by studying the change types surrounding these releases, I could get a better understanding of
what types of source code changes or trends constitute software stability or maturity. In order to study the release points, I went to each of our 10 project's 
web pages and looked through their release histories for major, minor, alpha, beta, and release candidate type releases. In total I identified 472 releases
across my 10 studied projects.

@@

Equation~\ref{eq:norm} works to create a change ratio by first creating a numerator by summing across all releases of
a given release type a sum of a particular change type in commits ($T_c$)
from the release date ($r$) to a given number of days after the release ($d$) divided by the number of commits in this date range ($|c|$). Next the denominator
is created by summing across all release of a given release type
a sum of that same particular change type in commits ($T_c$) from a given number of days ($d$) before the release date ($r$) to
the release date divided by the number of commits in this date range ($|c|$). This numerator and denominator form the final change ratio.
This equation gives us a ratio of a particular change type happening before and after a particular release. If the ratio is above 1 then that particular change
type occurs more frequently after the release and if it is below 1 then it occurs more frequently before the release. For the purposes of our study, we set the
number of days before and after the release ($d$) to 60 as the projects studies had many months in between their major releases. This quantitative data
formed much of the basis for the results to come in Section~\ref{sec:apie-results}

@@

Lastly, I found that software changes related to testing can be an indicator of a major release points within the projects studied. The change
ratios found can be seen in Table~\ref{tab:test}. As it can be seen, the four ratios which meet our threshold and are indicators of stability with regards to test based
changes are: the changing of test classes, the removal of test classes, the adding of methods, and the changing of method signatures, and test classes being changed.
Changes to documentation also meets my threshold and occurs more before a major release.

@@

In this study, I have conducted a case study of 10 open source Java software projects in order to study their change trends surrounding
major releases as previous studies have shown that indirect conflicts occur less at a major release and more so after a major release
or at the beginning of a development cycle. I have presented here 9 major change trends which surround major releases in the open source
projects studied. The 4 change trends found in this study occurring before major releases are: added private classes, 
changed test method signatures, changed documentation, and removed test classes.
The 5 change trends found in this paper occurring after major release are: added test methods, changed test classes, removed public methods, removed
private classes, and removed private methods.

>>>>

Once the ChangeDistiller information was collected, I decided to examine software change trends surrounding releases of the project's I had selected. Since releases
have preconceived notions of software stability and a lack of proneness to indirect conflicts, 
I decided that by studying the change types surrounding these releases, I could get a better understanding of
what types of source code changes or trends constitute software stability or maturity. In order to study the release points, I went to each of my 10 project's 
web pages and looked through their release histories for major, minor, alpha, beta, and release candidate type releases. In total I identified 472 releases
across my 10 studied projects.

@@

Equation~\ref{eq:norm} works to create a change ratio by first creating a numerator by summing across all releases of
a given release type a sum of a particular change type in commits ($T_c$)
from the release date ($r$) to a given number of days after the release ($d$) divided by the number of commits in this date range ($|c|$). Next the denominator
is created by summing across all release of a given release type
a sum of that same particular change type in commits ($T_c$) from a given number of days ($d$) before the release date ($r$) to
the release date divided by the number of commits in this date range ($|c|$). This numerator and denominator form the final change ratio.
This equation gives us a ratio of a particular change type happening before and after a particular release. If the ratio is above 1 then that particular change
type occurs more frequently after the release and if it is below 1 then it occurs more frequently before the release. For the purposes of my study, I set the
number of days before and after the release ($d$) to 60 as the projects studies had many months in between their major releases. This quantitative data
formed much of the basis for the results to come in Section~\ref{sec:apie-results}

@@

Lastly, I found that software changes related to testing can be an indicator of a major release points within the projects studied. The change
ratios found can be seen in Table~\ref{tab:test}. As it can be seen, the four ratios which meet my threshold and are indicators of stability with regards to test based
changes are: the changing of test classes, the removal of test classes, the adding of methods, and the changing of method signatures, and test classes being changed.
Changes to documentation also meets my threshold and occurs more before a major release.

@@

In this study, I have conducted a case study of 10 open source Java software projects in order to study their change trends surrounding
major releases as previous studies have shown that indirect conflicts occur less at a major release and more so after a major release
or at the beginning of a development cycle. I have presented here 9 major change trends which surround major releases in the open source
projects studied. The 4 change trends found in this study occurring before major releases are: added private classes, 
changed test method signatures, changed documentation, and removed test classes.
The 5 change trends found in this study occurring after major release are: added test methods, changed test classes, removed public methods, removed
private classes, and removed private methods.

### 2014-02-11 msc-thesis/chapters/3/sec-iced.tex
MSC danas edits.

In order to fully understand the root causes of information overload, false positives, and
scalability issues in regards to indirect conflicts, I conducted a large scale empirical study involving industrial
participants in order to determine: what events occur to
cause indirect conflicts, when they occur, if compounding conditions exist, and what mitigation strategies developers currently
use. Through this large exploration, what can be accomplished moving forward with indirect conflicts
in both research and industry and how we can get over the road block of these information overload related
issues. My research questions for this particular study are as follows:

\begin{description}
        \item[RQ1\namedlabel{itm:rq1}{RQ1}] \textit{What events or conditions lead to indirect conflicts?}
        \item[RQ2\namedlabel{itm:rq2}{RQ2}] \textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
        \item[RQ3\namedlabel{itm:rq3}{RQ3}] \textit{What do developers want from future indirect conflict tools?}
\end{description}

@@

Many indirect conflict tools have been created, tested, and published. Sarma et al.~\cite{Sarma:2007:TSA} created Palantir,
which can both detect potential indirect conflicts, at the class signature level, and alert developers to these conflicts.
Palantir represented one of the first serious attempts at aiding developers towards indirect conflicts. Holmes et
al.~\cite{Holmes:2010:CAR} take it one step further with their tool YooHoo, by detecting fine grained source code changes,
such as method return type changes, and create a taxonomy for different types of changes and their proneness to cause
indirect conflicts. This tool and its taxonomy had a severely reduced false positive rate, however, the true positives
detected may already be detectable by current tools such as compilers and unit testing. The tool Ariadne~\cite{Trainer:2005:BGT}
creates an environment where developers can see how source code changes will affect other areas of a project at the
method level, using method call graphs, showing where indirect conflicts may occur. This type of exploratory design has
been used often in the visualization of indirect conflict tools, allowing developers a type of search area for their development
needs. Another indirect conflict tool, CASI~\cite{Servant:2010:CPI}, utilizes dependency slicing~\cite{Bajracharya:2009:SIS}
instead of method call graphs to provide an environment to see what areas of a project are being affected by a source code change.
Most of these tools have all shown to have the same common difficulties of scalability, false positives, and information overload,
which were explored in this paper. Our own tool
Impact!~\footnote{University of Victoria Computer Science Technical Paper: DCS-351-IR} also suffered these same fates.

@@

For developer interest in regards to software modifications, Kim~\cite{Kim:2011:ESA} found that developers wanted to know whose
recent code changes semantically interfere with their own code changes, and whether their code is impacted by a code change.
Kim conducted several focus groups as well as web questionnaires to determine these results. These areas of interest resonate with what was
found in this paper. Developers want to know when a code change is going to interfere with their work in a potentially negative way and
require action on the developer's half.  Kim
also found that developers are concerned with interfaces of objects and when those interfaces change, similar to the object contracts that
were found in this paper. Finally, Kim also identified the same issues towards information overload through false positives with developers
noting ``I get a big laundry list... I see the email and I delete it''.

In a case study of impact awareness, de Souza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
code base to determine the impact of their code changes on fellow developers, teams, and projects. This corresponds with the findings of
preventative measures I have found in that some human interaction or knowledge is required in preventing indirect conflicts along with
development processes.

@@

Interview participants first answered a number of demographic questions. I then asked them to describe various software development experiences regarding our three research questions.
Specifically, ten semi-structured  topics directly related to our research questions guided our interview:

@@

My questionnaire respondents were different from my interviewees. I invited our questionnaire participants from a similar breadth of open and closed source software development 
companies and projects as the interviews participants with two main exceptions. The software organizations
that remained the same between interview and questionnaire were: Mozilla, The GNOME Project, Microsoft Corporation, 
Subnet Solutions, and Amazon.
Participants who took part in the round of interviews were not invited to the questionnaire but were asked to act as a point of contact for other 
developers in their team, project, or organization who may be interested
in completing the questionnaire. Further, two other groups of developers were asked to
participate as well, these being GitHub users as well as Apache Software Foundation (Apache) developers. The GitHub
users were selected based on large amounts of development activity on GitHub and the Apache developers
were selected based on their software development contributions on specific projects known to be used heavily
utilized by other organizations and projects.

@@

\subsubsection{What events or conditions lead to indirect conflicts?}

@@

11 developers explained that indirect conflicts occur ``all the time'' in
their development life cycle with a minimum occurrence of once a week, with more serious issues tending
to occur once a month. To confirm this, 64\% of questionnaire participants answered that indirect conflicts occur on minimum bi-weekly,
with the 25\% saying that weekly occurrences are most common (seen in ~\ref{tab:often}). 5 questionnaire participants
also stated that the stage of the development cycle can greatly cause the frequency of indirect conflicts to
differ.

12 developers said that
when a project is in the early stages of development, indirect conflicts tend to occur far more frequently
than once a stable point is reached. Developers said ``At a stable point we decided we are not going to change
[this feature] anymore. We will only add new code instead of changing it.'' and ``the beginning of a project
changes a lot, especially with agile''. Questionnaire participants also added ``indirect conflicts after a release
depend on how well the project was built at first'', ``[indirect conflicts] tend to slow down a bit after a
major release, unless the next release is a major rework.'', and ``[indirect conflicts have] spikes during
large revamps or the implementation of cross-cutting new features.'' in order to confirm our interview results.
Questionnaire participants also answered that
indirect conflicts are more likely to occur before the first major release rather than after at the daily
and weekly occurrence intervals as seen in Table~\ref{tab:often}.

@@

@@

From the re-interviewed participants, I found extremely positive feedback regarding both my results and major discussion
points. Participants often had new stories and experiences to share once they had heard the results of this paper which
confirmed the findings and often were quite shocked to hear the results as they did not usually think about their actions
as such but then realized the results held true to their daily actions for better or worse.

@@

{\bfseries Applicability or Usefulness.} ``Do the findings offer new insights? Can they be used to develop policy or change practice?''
Although my results may not be entirely novel or even surprising, the combination of said results allowed me to discover a
the disjoint between theoretical awareness and practical awareness regarding indirect conflict tools as well as provide more
insight into the debate of prevention versus cure in software development (as to be seen in Chapter~\ref{chapter:discussion}). 
Given how many indirect conflict tools are left with the same common
issues, I believe that these findings will help researchers focus on what developers want and need moving into the future more than has
been possible in the past. These finding set a course of action for where effort should be spent in academia to better benefit industry.

10 of the 78 questionnaire participants sent direct responses to me asking for any results of the research to be sent directly to
them in order to improve their indirect conflict work flows. 7 of the 19 participants interviewed expressed interest concerning any possible
tools or plans for tools inspired by this research as well. The combination of academia relatability and direct industry interest 
in my results help us fulfill this criterion.

@@

\subsection{Threats to Validity}
% TODO

@@

My findings indicate that: 
indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding of change impacts,
indirect conflicts occur quite frequently with a tendency to be more prevalent at the beginning of a development cycle, indirect conflicts
occur more often as a software project grows in developer numbers, developers use design by contract, add and deprecate, and experience
to mitigate indirect conflicts, and finally, developer prefer to have tools which help them debug quicker once a conflict has occurred rather
than tools which help prevent possible conflicts for the future.

>>>>

I want to understand why it is so hard to tackle indirect conflicts, specifically through tool based solutions.
To do so, I take a step back and intend to obtain a broader view of indirect conflicts
with a large field study of practitioners. I investigate the root causes of information overload, false positives, and
scalability issues in regards to indirect conflicts to better understand why indirect conflict tools fail to achieve
the success of other domain tools. I determine: general events which 
cause indirect conflicts, when they occur, if compounding conditions exist, mitigation strategies developers
use, and what practitioners want from new tools. My research questions for this particular study are as follows:

\begin{description}
        \item[RQ1\namedlabel{itm:rq1}{RQ1}] \textit{What are the types, factors, and frequencies of indirect conflicts?}
        \item[RQ2\namedlabel{itm:rq2}{RQ2}] \textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
        \item[RQ3\namedlabel{itm:rq3}{RQ3}] \textit{What do developers want from future indirect conflict tools?}
\end{description}

@@

Many indirect conflict tools have been created, tested, and published. Sarma et al.~\cite{Sarma:2007:TSA} created Palantir,
which can both detect potential indirect conflicts, at the class signature level, and alert developers to these conflicts.
Palantir represented one of the first serious attempts at aiding developers towards indirect conflicts. Holmes et
al.~\cite{Holmes:2010:CAR} take it one step further with their tool YooHoo, by detecting fine grained source code changes,
such as method return type changes, and create a taxonomy for different types of changes and their proneness to cause
indirect conflicts. This tool and its taxonomy had a severely reduced false positive rate, however, the true positives
detected may already be detectable by current tools such as compilers and unit testing. The tool Ariadne~\cite{Trainer:2005:BGT}
creates an environment where developers can see how source code changes will affect other areas of a project at the
method level, using method call graphs, showing where indirect conflicts may occur. This type of exploratory design has
been used often in the visualization of indirect conflict tools, allowing developers a type of search area for their development
needs. Another indirect conflict tool, CASI~\cite{Servant:2010:CPI}, utilizes dependency slicing~\cite{Bajracharya:2009:SIS}
instead of method call graphs to provide an environment to see what areas of a project are being affected by a source code change.
Most of these tools have all shown to have the same common difficulties of scalability, false positives, and information overload,
which were explored in this study. My own tool
Impact!~\cite{Ell:2013} also suffered these same fates.

@@

For developer interest in regards to software modifications, Kim~\cite{Kim:2011:ESA} found that developers wanted to know whose
recent code changes semantically interfere with their own code changes, and whether their code is impacted by a code change.
Kim found that developers are concerned with interfaces of objects and when those interfaces change. Kim also identified the same 
issues towards information overload through false positives with developers noting ``I get a big laundry list... I see the email and I delete it''.
Kim's field study does however fall short in actually creating a resolution for indirect conflicts, or finding new concerns of
developers which are not already detected by compilation or other static analysis tools. 
For impact awareness, DeSouza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
code base to determine the impact of their code changes on fellow developers, teams, and projects. However, DeSouza does not
study which types of changes (types, frequencies, compounding factors) developers should concern themselves with more in terms of using their personal knowledge
to stop indirect conflicts. DeSouza also does not study formal mitigation strategies, or resolutions of indirect conflicts directly.

This study is intended to fill the gap which has been left by aforementioned tools papers as well as the field study paper.
I will not only study why information overload, false positives, and scalability are such difficult problems to tackle
in indirect conflict tools, but I will also study how developers currently deal with indirect conflicts in practice through
their mitigation strategies, their largest concerns, and their ideas for future suggestions in regards to indirect conflicts.

@@

Interview participants first answered a number of demographic questions. I then asked them to describe various software development experiences regarding my three research questions.
Specifically, ten semi-structured topics directly related to my research questions guided my interview:

@@

My questionnaire respondents were different from my interviewees. I invited my questionnaire participants from a similar breadth of open and closed source software development 
companies and projects as the interviews participants with two main exceptions. The software organizations
that remained the same between interview and questionnaire were: Mozilla, The GNOME Project, Microsoft Corporation, 
Subnet Solutions, and Amazon.
Participants who took part in the round of interviews were not invited to the questionnaire but were asked to act as a point of contact for other 
developers in their team, project, or organization who may be interested
in completing the questionnaire. Further, two other groups of developers were asked to
participate as well, these being GitHub users as well as Apache Software Foundation (Apache) developers. The GitHub
users were selected based on large amounts of development activity on GitHub and the Apache developers
were selected based on their software development contributions on specific projects known to be used heavily
utilized by other organizations and projects.

@@

\subsubsection{What are the types, factors, and frequencies of indirect conflicts?}

\bf{The most common occurrence of indirect conflict is when a software object's contract changes. The frequency of
indirect conflicts, while usually high, decreases as a stable point is reached in the development cycle. The frequency
of indirect conflicts is compounded by the number of developers working on a project.}

\normalfont{}

@@

11 developers explained that indirect conflicts occur ``all the time'' in
their development life cycle with a minimum occurrence of once a week, with more serious issues tending
to occur once a month. To confirm this, 64\% of questionnaire participants answered that indirect conflicts occur bi-weekly or more
frequent,
with the 25\% saying that weekly occurrences are most common (seen in Table~\ref{tab:often}). 5 questionnaire participants
also stated that the stage of the development cycle can greatly cause the frequency of indirect conflicts to
differ.

12 developers said that
when a project is in the early stages of development, indirect conflicts tend to occur far more frequently
than once a stable point is reached. Developers said ``At a stable point we decided we are not going to change
[this feature] anymore. We will only add new code instead of changing it.'' and ``the beginning of a project
changes a lot, especially with agile''. Questionnaire participants also added ``indirect conflicts after a release
depend on how well the project was built at first'', ``[indirect conflicts] tend to slow down a bit after a
major release, unless the next release is a major rework.'', and ``[indirect conflicts have] spikes during
large revamps or the implementation of cross-cutting new features.'' in order to confirm mu interview results.
Questionnaire participants also answered that
indirect conflicts are more likely to occur before the first major release rather than after at the daily
and weekly occurrence intervals as seen in Table~\ref{tab:often}.

@@

\bf{Being that developers believe ``good enough'' solutions exist for preventing and detecting indirect conflicts
developers would rather see improvement to debugging processes. Developers are looking for automated debugging tools,
and better source code analysis tool for impact management and supporting cross team and cross sub-project projects.}

\normalfont{}

@@

From the re-interviewed participants, I found extremely positive feedback regarding both my results and major discussion
points. Participants often had new stories and experiences to share once they had heard the results of this study which
confirmed the findings and often were quite shocked to hear the results as they did not usually think about their actions
as such but then realized the results held true to their daily actions for better or worse.

@@

{\bfseries Applicability or Usefulness.} ``Do the findings offer new insights? Can they be used to develop policy or change practice?''
Although my results may not be entirely novel or even surprising, the combination of said results allowed me to discover a
the disjoint between theoretical awareness and practical awareness regarding indirect conflict tools as well as provide more
insight into the debate of prevention versus cure in software development (as to be seen in Chapter~\ref{chapter:discussion}). 
Given how many indirect conflict tools are left with the same common
issues, I believe that these findings will help researchers focus on what developers want and need moving into the future more than has
been possible in the past. These finding set a course of action for where effort should be spent in research to better benefit industry.

10 of the 78 questionnaire participants sent direct responses to me asking for any results of the research to be sent directly to
them in order to improve their indirect conflict work flows. 7 of the 19 participants interviewed expressed interest concerning any possible
tools or plans for tools inspired by this research as well. The combination of research relatability and direct industry interest 
in my results help us fulfill this criterion.

@@

\subsection{Threats to Validity}
Because of the exploratory nature of this work, I chose Grounded Theory as my research method, which has some implications regarding the
limitations and threats to validity of this study. While I achieved high saturation regarding the topics focused on in this study, other
populations studied from outside of this studies participants may add new insights into the pool of findings. As a result of this, findings
from this study may not relate to everyone or generalize to outside of the group of participants involved in this study.

As per my interview participants, I attempted to cover as great a breadth as possible of software developers in their practices, companies,
project, languages, etc. However, it would be extremely difficult to conduct a completely thorough study of indirect conflicts in practice
as there exist too many demographic options for participants. This same applies for questionnaire participants. This being said, questionnaire 
or interview results could become heavily biased towards a developers good or bad experiences with any given project.

To counter act this, my confirmatory interviews found high agreement for several of my findings. Therefore, I believe that
I have uncovered valuable insight regarding indirect conflicts in practice.

@@

My findings have indicated that:
indirect conflicts occur frequently and are likely caused by software contract changes;
while design by contract, add and deprecate, and personal experience help prevent indirect conflicts,
developers tend to prefer to use detection and resolution processes or tools
over those of prevention, and developers want indirect conflict tools to focus on automatic debugging and better source code analysis.

My result analysis has indicated that: developers do not want awareness mechanisms which provide non actionable results; developers
would rather focus on curing existing problems rather than preventing potential issues;
and that there exists a gap in software evolution analytical tools between what is available and what practitioners need.
